---
title: "<center>Final Project Data Cleaing</center>"
author: "<center>111071033</center>"
output:
  html_document: 
    theme: darkly
    highlight: monochrome
    df_print: paged
    toc: true
    toc_float: true
    css: style.css
  html_notebook: 
    theme: darkly
    highlight: monochrome
    df_print: paged
    toc: true
    toc_float: true
    css: style.css
    code_folding: show
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = FALSE,
                      comment = "",
                      fig.height = 5,
                      fig.width = 10.4,
                      class.output = ".codeout")
#theme_set(theme_bw())
```
|
|
## Prepare Requisite

### load package
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(stringr)
library(tidyr)
library(purrr)
```


### load file
```{r}
revenue_raw <- read_csv("lysk_iphone_revenue_20240118_20260615.csv")
events_raw  <- read_excel("lysk_pull_event_info.xlsx")
```


## data cleaning
### clean daily revenue data
```{r}
revenue_daily_clean <- revenue_raw |> 
  transmute(
    date = ymd(`日期`),
    revenue_ios = as.numeric(`收入预估~iPhone`)
  ) |> 
  filter(!is.na(date), !is.na(revenue_ios)) |> 
  arrange(date) |> 
  mutate(
    log_revenue_ios = log(revenue_ios + 1),
    year = year(date),
    month = month(date),
    week = isoweek(date),
    weekday = wday(date, label = TRUE, week_start = 1),
    is_weekend = weekday %in% c("Sat", "Sun","周六","周日")
  )

glimpse(revenue_daily_clean)
summary(revenue_daily_clean$revenue_ios)
```

### clean pull event data
```{r}
events_clean <- events_raw |> 
  transmute(
    event_id = row_number(),
    event_name = event_name,
    pv_upload_date = ymd(pv_upload_date),
    event_start = ymd(event_start),
    event_end = ymd(event_end),
    type = type,
    pool_scope = `single/all`,
    is_return = is_return,
    character_include = character_include,
    special_holiday = special_holiday,
    public_opinion = public_opinion
  ) |> 
  mutate(
    event_duration_days = as.integer(event_end - event_start) + 1,
    pv_to_start_days = as.integer(event_start - pv_upload_date),

    is_moon = type == "moon",
    is_sun = type == "sun",
    is_roulette = type == "4-star_roulutte",

    is_single = pool_scope == "single",
    is_all = pool_scope == "all",

    is_return_dummy = is_return == "Y",

    # 統一角色分隔符
    character_include_std = character_include |> 
      str_replace_all("、|，", ",") |> 
      str_replace_all("\\s+", ""),

    n_characters = if_else(
      is.na(character_include_std) | character_include_std == "",
      0L,
      str_count(character_include_std, ",") + 1L
    )
  )

glimpse(events_clean)
```

### seperate character_include_std into dummy columns
```{r}
events_clean <- events_clean |> 
  mutate(
    has_shenxinghui = str_detect(coalesce(character_include_std, ""), "沈星回"),
    has_lishen = str_detect(coalesce(character_include_std, ""), "黎深"),
    has_qiyu = str_detect(coalesce(character_include_std, ""), "祁煜"),
    has_qinche = str_detect(coalesce(character_include_std, ""), "秦徹"),
    has_xiayizhou = str_detect(coalesce(character_include_std, ""), "夏以晝")
  )
```

### seperate and label special_holiday
```{r}
events_clean <- events_clean |> 
  mutate(
    special_holiday = replace_na(special_holiday, "N"),

    is_birthday = str_detect(special_holiday, "角色生日|生日"),

    is_special_holiday = str_detect(
      special_holiday,
      "情人節|白色情人節|五一|國慶|農曆新年|七夕"
    ),

    is_anniversary = str_detect(
      special_holiday,
      "一周年|二周年|周年"
    ),

    is_new_version = str_detect(
      special_holiday,
      "版本更新|新增男主|開服"
    ),

    holiday_label = case_when(
      is_birthday ~ "birthday",
      is_anniversary ~ "anniversary",
      is_new_version ~ "new_version",
      is_special_holiday ~ "special_holiday",
      TRUE ~ "none"
    )
  )

events_clean |> 
  count(special_holiday, holiday_label, sort = TRUE)
```

### seperate and label public_opinion
```{r}
events_clean <- events_clean |> 
  mutate(
    public_opinion = replace_na(public_opinion, "none"),

    opinion_none = public_opinion == "none",

    opinion_negative = str_detect(
      public_opinion,
      "停氪維權|洩憤|負面|炎上|罵|不滿"
    ),

    opinion_positive_hype = str_detect(
      public_opinion,
      "狂熱|高討論|爆|熱度|出圈"
    ),

    opinion_external_conflict = str_detect(
      public_opinion,
      "rap圈|塵白禁區|外部風波|圈外"
    ),

    opinion_controversy = str_detect(
      public_opinion,
      "爭議|尺度|主題爭議|風波"
    )
  )
```


### label new character(male_lead) debut period
```{r}
qinche_first_date <- events_clean |> 
  filter(event_name == "不设防禁区") |> 
  pull(event_start) |> 
  first()

xiayizhou_first_date <- events_clean |> 
  filter(event_name == "触痛讯号") |> 
  pull(event_start) |> 
  first()

events_clean <- events_clean |> 
  mutate(
    is_qinche_fresh_period =
      is_single &
      has_qinche &
      event_start >= qinche_first_date &
      event_start < qinche_first_date %m+% months(3),

    is_xiayizhou_fresh_period =
      is_single &
      has_xiayizhou &
      event_start >= xiayizhou_first_date &
      event_start < xiayizhou_first_date %m+% months(3),

    is_new_character =
      is_qinche_fresh_period | is_xiayizhou_fresh_period,

    days_since_qinche_debut = if_else(
      has_qinche,
      as.integer(event_start - qinche_first_date),
      NA_integer_
    ),

    days_since_xiayizhou_debut = if_else(
      has_xiayizhou,
      as.integer(event_start - xiayizhou_first_date),
      NA_integer_
    )
  )
```

### exclude first event "心动变曲"
```{r}
first_xindong_date <- events_clean |> 
  filter(event_name == "心动变曲") |> 
  summarise(first_date = min(event_start, na.rm = TRUE)) |> 
  pull(first_date)

events_clean <- events_clean |> 
  mutate(
    is_first_launch_event =
      event_name == "心动变曲" & event_start == first_xindong_date,

    # 主分析只排首發心动变曲
    exclude_main_analysis = is_first_launch_event,

    # 核心主池分析再排 return
    exclude_core_pool_analysis = is_first_launch_event | is_return_dummy
  )

events_clean |> 
  select(
    event_name, event_start, is_first_launch_event,
    is_roulette, is_return_dummy,
    exclude_main_analysis, exclude_core_pool_analysis
  ) |> 
  arrange(event_start)
```

### event-day panel
```{r}
event_day_panel <- events_clean |> 
  mutate(date = map2(event_start, event_end, seq.Date, by = "day")) |> 
  unnest(date) |> 
  mutate(
    event_day = as.integer(date - event_start),
    is_active = TRUE,
    is_event_start_day = event_day == 0,
    is_event_day_0_2 = event_day %in% 0:2,
    is_event_day_0_6 = event_day %in% 0:6
  )
```


### daily event exposure
```{r}
daily_event_exposure <- event_day_panel |> 
  group_by(date) |> 
  summarise(
    active_event_count = n(),

    active_moon_count = sum(is_moon, na.rm = TRUE),
    active_sun_count = sum(is_sun, na.rm = TRUE),
    active_roulette_count = sum(is_roulette, na.rm = TRUE),
    active_return_count = sum(is_return_dummy, na.rm = TRUE),

    has_moon = active_moon_count > 0,
    has_sun = active_sun_count > 0,
    has_roulette = active_roulette_count > 0,
    has_return = active_return_count > 0,

    has_single = sum(is_single, na.rm = TRUE) > 0,
    has_all = sum(is_all, na.rm = TRUE) > 0,

    is_any_event_start = sum(is_event_start_day, na.rm = TRUE) > 0,
    is_moon_start_day = sum(is_event_start_day & is_moon, na.rm = TRUE) > 0,
    is_sun_start_day = sum(is_event_start_day & is_sun, na.rm = TRUE) > 0,
    is_roulette_start_day = sum(is_event_start_day & is_roulette, na.rm = TRUE) > 0,
    is_return_start_day = sum(is_event_start_day & is_return_dummy, na.rm = TRUE) > 0,

    active_shenxinghui_count = sum(has_shenxinghui, na.rm = TRUE),
    active_lishen_count = sum(has_lishen, na.rm = TRUE),
    active_qiyu_count = sum(has_qiyu, na.rm = TRUE),
    active_qinche_count = sum(has_qinche, na.rm = TRUE),
    active_xiayizhou_count = sum(has_xiayizhou, na.rm = TRUE),

    active_birthday_count = sum(is_birthday, na.rm = TRUE),
    active_special_holiday_count = sum(is_special_holiday, na.rm = TRUE),
    active_anniversary_count = sum(is_anniversary, na.rm = TRUE),
    active_new_version_count = sum(is_new_version, na.rm = TRUE),

    active_negative_opinion_count = sum(opinion_negative, na.rm = TRUE),
    active_positive_hype_count = sum(opinion_positive_hype, na.rm = TRUE),
    active_external_conflict_count = sum(opinion_external_conflict, na.rm = TRUE),
    active_controversy_count = sum(opinion_controversy, na.rm = TRUE),

    .groups = "drop"
  )
```

### daily_panel
```{r}
daily_panel <- revenue_daily_clean |> 
  left_join(daily_event_exposure, by = "date") |> 
  mutate(
    across(
      c(
        active_event_count,
        active_moon_count,
        active_sun_count,
        active_roulette_count,
        active_return_count,
        active_shenxinghui_count,
        active_lishen_count,
        active_qiyu_count,
        active_qinche_count,
        active_xiayizhou_count,
        active_birthday_count,
        active_special_holiday_count,
        active_anniversary_count,
        active_new_version_count,
        active_negative_opinion_count,
        active_positive_hype_count,
        active_external_conflict_count,
        active_controversy_count
      ),
      ~ replace_na(., 0)
    ),
    across(
      c(
        has_moon, has_sun, has_roulette, has_return,
        has_single, has_all,
        is_any_event_start, is_moon_start_day, is_sun_start_day,
        is_roulette_start_day, is_return_start_day
      ),
      ~ replace_na(., FALSE)
    )
  )
```

### event outcomes(general baseline)
```{r}
event_outcomes <- events_clean |> 
  rowwise() |> 
  mutate(
    rev_d1 = revenue_daily_clean |> 
      filter(date == event_start) |> 
      pull(revenue_ios) |> 
      first(),

    rev_d3_sum = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(2)) |> 
      summarise(x = sum(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    rev_d7_sum = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(6)) |> 
      summarise(x = sum(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    rev_avg_d3 = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(2)) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    rev_avg_d7 = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(6)) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    baseline_pre3 = revenue_daily_clean |> 
      filter(date >= event_start - days(3), date <= event_start - days(1)) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    baseline_pre7 = revenue_daily_clean |> 
      filter(date >= event_start - days(7), date <= event_start - days(1)) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x)
  ) |> 
  ungroup() |> 
  mutate(
    abnormal_rev_d1 = rev_d1 - baseline_pre3,
    abnormal_rev_d3_sum = rev_d3_sum - 3 * baseline_pre3,
    abnormal_rev_d7_sum = rev_d7_sum - 7 * baseline_pre7,

    log_rev_d1 = log(rev_d1 + 1),
    log_rev_d3_sum = log(rev_d3_sum + 1),

    log_abnormal_rev_d1 = log(pmax(abnormal_rev_d1, 0) + 1),
    log_abnormal_rev_d3_sum = log(pmax(abnormal_rev_d3_sum, 0) + 1)
  )
```


### clean baseline excluding "心动变曲"
```{r}
excluded_dates_for_baseline <- event_day_panel |> 
  filter(is_first_launch_event) |> 
  distinct(date) |> 
  mutate(is_excluded_baseline_date = TRUE)

revenue_daily_for_baseline <- revenue_daily_clean |> 
  left_join(excluded_dates_for_baseline, by = "date") |> 
  mutate(
    is_excluded_baseline_date = replace_na(is_excluded_baseline_date, FALSE)
  )
```

### event outcomes(clean baseline excluding "心动变曲")
```{r}
event_outcomes_clean_baseline <- events_clean |> 
  rowwise() |> 
  mutate(
    rev_d1 = revenue_daily_clean |> 
      filter(date == event_start) |> 
      pull(revenue_ios) |> 
      first(),

    rev_d3_sum = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(2)) |> 
      summarise(x = sum(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    rev_d7_sum = revenue_daily_clean |> 
      filter(date >= event_start, date <= event_start + days(6)) |> 
      summarise(x = sum(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    clean_pre3_n = revenue_daily_for_baseline |> 
      filter(
        date >= event_start - days(3),
        date <= event_start - days(1),
        !is_excluded_baseline_date
      ) |> 
      nrow(),

    clean_pre7_n = revenue_daily_for_baseline |> 
      filter(
        date >= event_start - days(7),
        date <= event_start - days(1),
        !is_excluded_baseline_date
      ) |> 
      nrow(),

    baseline_pre3_clean = revenue_daily_for_baseline |> 
      filter(
        date >= event_start - days(3),
        date <= event_start - days(1),
        !is_excluded_baseline_date
      ) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x),

    baseline_pre7_clean = revenue_daily_for_baseline |> 
      filter(
        date >= event_start - days(7),
        date <= event_start - days(1),
        !is_excluded_baseline_date
      ) |> 
      summarise(x = mean(revenue_ios, na.rm = TRUE)) |> 
      pull(x)
  ) |> 
  ungroup() |> 
  mutate(
    abnormal_rev_d1_clean = rev_d1 - baseline_pre3_clean,
    abnormal_rev_d3_sum_clean = rev_d3_sum - 3 * baseline_pre3_clean,
    abnormal_rev_d7_sum_clean = rev_d7_sum - 7 * baseline_pre7_clean,

    log_abnormal_rev_d1_clean = log(pmax(abnormal_rev_d1_clean, 0) + 1),
    log_abnormal_rev_d3_sum_clean = log(pmax(abnormal_rev_d3_sum_clean, 0) + 1),

    baseline_clean_fallback = case_when(
      !is.nan(baseline_pre3_clean) ~ baseline_pre3_clean,
      !is.nan(baseline_pre7_clean) ~ baseline_pre7_clean,
      TRUE ~ NA_real_
    ),

    abnormal_rev_d1_clean_fb = rev_d1 - baseline_clean_fallback,
    abnormal_rev_d3_sum_clean_fb = rev_d3_sum - 3 * baseline_clean_fallback,

    log_abnormal_rev_d1_clean_fb = log(pmax(abnormal_rev_d1_clean_fb, 0) + 1),
    log_abnormal_rev_d3_sum_clean_fb = log(pmax(abnormal_rev_d3_sum_clean_fb, 0) + 1)
  ) |> 
  select(
    event_id,
    clean_pre3_n,
    clean_pre7_n,
    baseline_pre3_clean,
    baseline_pre7_clean,
    abnormal_rev_d1_clean,
    abnormal_rev_d3_sum_clean,
    abnormal_rev_d7_sum_clean,
    log_abnormal_rev_d1_clean,
    log_abnormal_rev_d3_sum_clean,
    baseline_clean_fallback,
    abnormal_rev_d1_clean_fb,
    abnormal_rev_d3_sum_clean_fb,
    log_abnormal_rev_d1_clean_fb,
    log_abnormal_rev_d3_sum_clean_fb
  )
```

### merge clean baseline back to event_outcomes
```{r}
event_outcomes <- event_outcomes |> 
  select(
    -any_of(c(
      "clean_pre3_n",
      "clean_pre7_n",
      "baseline_pre3_clean",
      "baseline_pre7_clean",
      "abnormal_rev_d1_clean",
      "abnormal_rev_d3_sum_clean",
      "abnormal_rev_d7_sum_clean",
      "log_abnormal_rev_d1_clean",
      "log_abnormal_rev_d3_sum_clean",
      "baseline_clean_fallback",
      "abnormal_rev_d1_clean_fb",
      "abnormal_rev_d3_sum_clean_fb",
      "log_abnormal_rev_d1_clean_fb",
      "log_abnormal_rev_d3_sum_clean_fb"
    ))
  ) |> 
  left_join(event_outcomes_clean_baseline, by = "event_id")
```


### D1–D3 overlap
```{r}
event_windows_d3 <- events_clean |> 
  transmute(
    event_id,
    event_name,
    window_date = map(event_start, ~ seq(.x, .x + days(2), by = "day"))
  ) |> 
  unnest(window_date)
```

### overlap controls
```{r}
overlap_raw <- event_windows_d3 |> 
  left_join(
    event_day_panel |> 
      transmute(
        event_id_other = event_id,
        date,
        is_return_dummy,
        is_roulette
      ),
    by = c("window_date" = "date"),
    relationship = "many-to-many"
  )

overlap_info <- overlap_raw |> 
  filter(event_id != event_id_other) |> 
  mutate(
    is_return_only = is_return_dummy & !is_roulette,
    is_roulette_only = is_roulette & !is_return_dummy,
    is_return_and_roulette = is_return_dummy & is_roulette
  ) |> 
  group_by(event_id) |> 
  summarise(
    overlap_days_d3 = n_distinct(window_date),
    overlap_event_rows_d3 = n(),
    overlap_return_rows_d3 = sum(is_return_dummy, na.rm = TRUE),
    overlap_roulette_rows_d3 = sum(is_roulette, na.rm = TRUE),
    overlap_return_only_rows_d3 = sum(is_return_only, na.rm = TRUE),
    overlap_roulette_only_rows_d3 = sum(is_roulette_only, na.rm = TRUE),
    overlap_return_and_roulette_rows_d3 = sum(is_return_and_roulette, na.rm = TRUE),
    .groups = "drop"
  )
```

### merge overlap info back to event_outcomes
```{r}
event_outcomes <- event_outcomes |> 
  select(-starts_with("overlap_")) |> 
  left_join(overlap_info, by = "event_id") |> 
  mutate(
    overlap_days_d3 = replace_na(overlap_days_d3, 0),
    overlap_event_rows_d3 = replace_na(overlap_event_rows_d3, 0),
    overlap_return_rows_d3 = replace_na(overlap_return_rows_d3, 0),
    overlap_roulette_rows_d3 = replace_na(overlap_roulette_rows_d3, 0),
    overlap_return_only_rows_d3 = replace_na(overlap_return_only_rows_d3, 0),
    overlap_roulette_only_rows_d3 = replace_na(overlap_roulette_only_rows_d3, 0),
    overlap_return_and_roulette_rows_d3 = replace_na(overlap_return_and_roulette_rows_d3, 0)
  )
```

### 3 slice of sample
```{r}
event_outcomes_all <- event_outcomes

event_outcomes_main <- event_outcomes |> 
  filter(!exclude_main_analysis)

event_outcomes_core <- event_outcomes |> 
  filter(!exclude_core_pool_analysis)

sample_overview <- tibble(
  sample_name = c("all", "main", "core"),
  n = c(
    nrow(event_outcomes_all),
    nrow(event_outcomes_main),
    nrow(event_outcomes_core)
  )
)


sample_overview
```

### save cleaned outcomes
```{r}
write_csv(revenue_daily_clean, "revenue_daily_clean.csv")
write_csv(events_clean, "events_clean.csv")
write_csv(event_day_panel, "event_day_panel.csv")
write_csv(daily_panel, "daily_panel.csv")
write_csv(event_outcomes, "event_outcomes.csv")
write_csv(event_outcomes_all, "event_outcomes_all.csv")
write_csv(event_outcomes_main, "event_outcomes_main.csv")
write_csv(event_outcomes_core, "event_outcomes_core.csv")
```

